On the use of petri-nets for modeling concurrent programs

### 2010-07-06 Documents/Publications/type-2/type-2.tex
modified the abstract..

\title{A class of untimed live and safe petri-nets that supports
  scalable static analysis}

@@

  Petri-nets are a compact way of representing concurrent sequences of
  events in a system, where events are represented as transitions
  connected to places that encode the state of the system. We present
  a class of untimed live and safe petri-nets called ``Type-2'' that
  allows a scalable static analysis, without severely affecting the
  expressive power of the petri-net.

  The simplest Type-2 petri-net is a cycle consisting of a single
  marked place and a single transition. Larger petri-nets are produced
  by progressive application of production rules that replace an
  unmarked place or a transition with a specific petri-net subgraph
  with no marked places. The initial marking of a Type-2 petri-net is
  thus restricted to the single marked place in the initial cycle. The
  production rules also ensure that every Type-2 petri-net is live and
  safe.

  The restriction on the structure of the Type-2 petri-net
  allows the definition of a simple labelling scheme that effectively
  simulates an execution of the petri-net. The labels of any two
  transitions can be compared in linear time to statically determine
  their relation in the execution of the petri-net. For example, the
  labels can be used determine whether two transitions can fire
  concurrently, thus identifying concurrent events in the system. It
  is also possible to generate an exhaustive listing of sets of
  transitions such that no two transitions in each set can fire
  simultaneously.

  Despite the restrictions imposed by the production rules, the Type-2
  class of petri-nets is powerful enough to represent synchronous
  systems. It has been used as a transition step in translating
  imperative programs to hardware descriptions\cite{ahir-thesis}. The
  resulting high-level synthesis flow uses the labelling scheme to
  optimise the generated hardware by identifying opportunities for
  sharing hardware resources in the absence of timing information.

>>>>

\title{On the use of petri-nets for modeling concurrent programs}

@@

  Petri-nets provide a powerful mechanism for the modeling of
  concurrent systems.   However, because of their generality, 
  it is often difficulty to prove properties about petri-net
  models unless some restrictions are placed on their construction.
  We consider the use of petri-nets to model algorithms described
  in high level programming languages.  In particular, we introduce
  a class of live and safe petri-nets (which we term as {\em Type-2}
  petri-nets) for this purpose.   This class is defined by a
  set of production rules which ensure liveness and safety of
  the resulting petri-net.  
  Further, we show that concurrency
  analysis of Type-2 petri-nets can be carried out very efficiently
  using a labelling scheme which is analogous to a symbolic simulation
  of the petri-net.
  Thus, it is easy to determine the sets of events (or transitions)
  which satisfy ordering relationships (two transitions are said to
  be ordered if it is impossible for both of them to fire simultaneously).  
  Such information is useful in identifying possibilities for conflict 
  free resource sharing in the implementations of the programs being
  modeled.
  Despite the restrictions imposed by the production rules, the Type-2
  class of petri-nets is powerful enough to represent programs
  written in traditional imperative languages such as C/C++ as well in synchronous
  languages such as Esterel. It has been used as a transition step in translating
  imperative programs to hardware descriptions\cite{ahir-thesis}. The
  resulting high-level synthesis flow uses the labelling scheme to
  optimise the generated hardware by identifying opportunities for
  sharing hardware resources in the absence of timing information.
